---
title: "wbb.qmd"
format: html
editor: visual
---

Lesson on Simulations chapter 14 in the book.

```{r}
library(tidyverse)
install.packages("ggbeeswarm")
library(ggbeeswarm)
library(ggrepel)
library(ggalt)
set.seed(1234)
```

Simulations allow us to say what we expect distribution of results to be

This says hey run me some simulations and use rbinomial. N is the number of simulations. size is the the number of something happening and the prob is the percentage. This says 1000 simulations, doing 20 shows, how many times will she make 5

```{r}
set.seed(1234)

simulations <- rbinom(n = 1000, size = 20, prob = .317)

table(simulations)

#top line is the number of makes. the bottom line is the number of times that outcome happens.
#The changes of 13 or more are next to 0. 
```

We think about this in terms of how often something happens or how unlikely something is to happen.

Cold Streaks

```{r}

set.seed(1234)

simulations <- rbinom(n = 1000, size = 15, prob = .326)

hist(simulations)



table(simulations)
```

streaks across time

the lead functions allows you to look at the next record.

```{r}

library(dplyr)

set.seed(1234)

simulations <- rbinom(n = 1000, size = 1, prob = .377)

four_in_a_row <- mean(simulations == 1 & lead(simulations, 1) == 1 & lead(simulations, 2) == 1 & lead(simulations, 3) == 1)

odds <- 1/four_in_a_row
```

stewie's percetnage .465 It would happen 10 times over 1000 where she would shoot 3-17

```{r}
set.seed(1234)

simulations <- rbinom(n = 1000, size = 17, prob = .465)

table(simulations)
```

Beesworm plots form Oct. 25 lesson

```{r}
players <- read_csv("https://thescoop.org/sports-data-files/wbb_players_2023.csv")
```

```{r}
activeplayers <- players |> filter( mp > 0, position != "NULL") 
```

```{r}
activeplayers <- activeplayers |> mutate(tspct=pts/(2*(fga+0.44*fta)))
```

focusing on players who shoot a lot

Summary calculates quartiles

```{r}
summary(activeplayers$fga)
```

```{r}
shooters <- activeplayers |> filter(fga > 187)
```

making the beesworm plot

```{r}
ggplot() + geom_beeswarm(data=shooters, aes(x=position, y=tspct), color="grey")
```

```{r}
umd <- activeplayers |> 
  filter(team == "Maryland") |> 
  filter(fga>187) |> 
  arrange(desc(tspct))
```

```{r}
ggplot() + 
  geom_beeswarm(
    data=shooters, 
    groupOnX=TRUE, 
    aes(x=position, y=tspct), color="grey") + 
  geom_beeswarm(
    data=umd, 
    groupOnX=TRUE, 
    aes(x=position, y=tspct), color="red")
```

```{r}
ggplot() + 
  geom_beeswarm(
    data=shooters, 
    groupOnX=TRUE, 
    aes(x=position, y=tspct), color="grey") + 
  geom_beeswarm(
    data=umd, 
    groupOnX=TRUE, 
    aes(x=position, y=tspct), color="red") +
  geom_text_repel(
    data=umd, 
    aes(x=position, y=tspct, label=player))
```

geom quasi random

```{r}
ggplot() + 
  geom_quasirandom(
    data=shooters, 
    groupOnX=TRUE, 
    aes(x=position, y=tspct), color="grey") + 
  geom_quasirandom(
    data=umd, 
    groupOnX=TRUE, 
    aes(x=position, y=tspct), color="red") +
  geom_text_repel(
    data=umd, 
    aes(x=position, y=tspct, label=player))
```

using geom_jitter

```{r}
ggplot() + 
  geom_jitter(
    data=shooters, 
    aes(x=position, y=tspct), color="grey") + 
  geom_jitter(
    data=umd, 
    aes(x=position, y=tspct), color="red") +
  geom_text_repel(
    data=umd, 
    aes(x=position, y=tspct, label=player))
```

Class from Nov. 30 using text annotations with graphs

```{r}
players <- read_csv("https://raw.githubusercontent.com/dwillis/hhs-snapshots/main/data/player_totals_20231130.csv") %>% 
  filter(mp >0 )

ggplot() + geom_point(data=players, aes(x=mp, y=pts))

topscorers <- players |> filter(pts > 175)

ggplot() + 
  geom_point(data=players, aes(x=mp, y=pts), color="grey") +
  geom_point(data=topscorers, aes(x=mp, y=pts), color="black")

ggplot() + 
  geom_point(data=players, aes(x=mp, y=pts), color="grey") + 
  geom_point(data=topscorers, aes(x=mp, y=pts), color="black") + 
  geom_encircle(data=topscorers, aes(x=mp, y=pts), s_shape=0, expand=0, colour="red")

ggplot() + 
  geom_point(data=players, aes(x=mp, y=pts), color="grey") + 
  geom_point(data=topscorers, aes(x=mp, y=pts), color="black") + 
  geom_text(data=topscorers, aes(x=mp, y=pts, label=full_name), hjust = 0, vjust=1) +
  geom_encircle(data=topscorers, aes(x=mp, y=pts), s_shape=.5, expand=.03, colour="red")

plot <- ggplot() + 
  geom_point(data=players, aes(x=mp, y=pts), color="grey") + 
  geom_point(data=topscorers, aes(x=mp, y=pts), color="black") +
  geom_point(data=clark, aes(x=mp, y=pts), color="yellow") +
  geom_encircle(data=topscorers, aes(x=mp, y=pts), s_shape=.5, expand=.03, colour="red") +
  geom_text(aes(x=275, y=275, label="Top scorers")) + 
  labs(title="Caitlin Clark Alone At Top", subtitle="The Iowa star is by far the top scorer among all NCAA players", x="Minutes", y="Points") + 
  theme_minimal() + 
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10), 
    panel.grid.minor = element_blank()
    )

clark <- topscorers %>% 
  filter(full_name == "Caitlin Clark")

```


---- Parity assignment ---------

```{r}
logs <- read_csv("https://thescoop.org/sports-data-files/wbblogs24.csv")
rankings <- read_csv("https://thescoop.org/sports-data-files/wbb_rankings.csv")
logs23 <- read_csv("https://thescoop.org/sports-data-files/wbblogs23.csv")
```

# Partity index score = (SRS)
# advantage of building an index is that there are a smaller range of numbers. 
```{r}
parity_index24 <- logs |> 
  group_by(Team, Conference) |> 
  summarise(srs_score = mean(TeamSRS), sos_score = mean(TeamSOS)) |> 
  mutate(parity_index = (srs_score + 100) / (sos_score + 100))

parity_index23 <- logs23 |> 
  group_by(Team, Conference) |> 
  summarise(srs_score = mean(TeamSRS), sos_score = mean(TeamSOS)) |> 
  mutate(parity_index = (srs_score + 100) / (sos_score + 100))
```

# scatter plot of b10
```{r}
parity_index24 |> 
  filter(Conference == 'Big Ten WBB' ) |> 
  ggplot() + 
  geom_point(aes(x=srs_score, y = sos_score, label = Team)) +
  geom_text(aes(x=srs_score, y = sos_score, label = Team))
```

# left joining
```{r}
parity_with_top25 <- parity_index24 |> left_join(rankings, join_by(Team == team))
```

Is there more parity this year than last year?
```{r}

combined_data <- bind_rows(parity_index24 |> mutate(season="2024"), parity_index23 |> mutate(season = "2023")) 

result <- combined_data %>%
  group_by(Team) %>%
  summarize(
    Parity_Index_2024 = mean(parity_index[season == "2024"]),
    Parity_Index_2023 = mean(parity_index[season == "2023"])
  ) %>%
  filter(!is.na(Parity_Index_2024)) |> 
  filter(!is.na(Parity_Index_2023)) |>
  ungroup() %>%
  summarise(
    p_value = t.test(Parity_Index_2024, Parity_Index_2023, paired = TRUE)$p.value,
    mean_difference = mean(Parity_Index_2024 - Parity_Index_2023)
  )

```