Volleyball assignment
Dylan Manfre

```{r}
library(tidyverse)
library(ggalt)
```

```{r}
#loading data
teams <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensVolleyballData/main/data/ncaa_womens_volleyball_matchstats_2023.csv")
```

Establishing averages based on metrics that I choose.

```{r}

#attempt 1
team_avgs <- teams %>% 
  group_by(team) %>% 
  summarize(
            killsperset = kills/s,
            acesperset = aces/s,
            digsperset = digs/s,
            def_kill_perset = defensive_kills/s,
            error_perset = errors/s,
            hit_perc_perset = hit_pct/s,
            service_error_perset = s_err/s) %>% 
  select(team, killsperset, acesperset, digsperset, def_kill_perset, error_perset, hit_perc_perset, service_error_perset)

view(team_avgs)

```
When running: lead teams first then do attempt 1 and then jump down to vbzscores.... on the second attempt, start over and run attempt 2 and then jump to vbzscores. 


Attempting to fix the issue with teams

My prompt for ChatGPT: When I run this code team_avgs <- teams %>% 
  group_by(team) %>% 
  summarize(
            killsperset = kills/s,
            acesperset = aces/s,
            digsperset = digs/s,
            def_kill_perset = defensive_kills/s,
            error_perset = errors/s,
            hit_perc_perset = hit_pct/s,
            service_error_perset = s_err/s) %>% 
  select(team, killsperset, acesperset, digsperset, def_kill_perset, error_perset, hit_perc_perset, service_error_perset) I still see multiple teams in the team column. The team is next to a date column which has each match from the 2021 volleyball season. I want to only see one team per row. when I run this code. What am I doing wrong and show me the code to fix this

Establishing averages based on metrics that I choose. 
```{r}
#attempt 2 using ChatGPT's code this time with the sums

team_avgs <- teams %>% 
  group_by(team) %>%
  summarize(
    killsperset = sum(kills) / sum(s),
    digsperset = sum(digs) / sum(s),
    hit_perc_perset = sum(hit_pct) / sum(s))

view(team_avgs)

#This now works and shows me each team as a single line which is what I wanted originally. The differnece is that it takes sum(kills)/sum(sets) which is different from how the tutorial showed it. 

```

creating the z scores
```{r}
vbzscore <- team_avgs %>%
  mutate(
    killszscore = as.numeric(scale(killsperset, center = TRUE, scale = TRUE)),
    digszscore = as.numeric(scale(digsperset, center = TRUE, scale = TRUE)),
    hit_perc_zscore = as.numeric(scale(hit_perc_perset, center = TRUE, scale = TRUE)),
    TotalZscore = killszscore + digszscore + hit_perc_zscore)


view(vbzscore)
```

class thoughts on the above code
- Teams being so high up 


it doesn't seem like the group_by worked because all the team names are still showing. Is this supposed to happen?

#First attempt at running this produces the zscore columns but they were all populated with NaNs
#Pasted this prompt to ChatGPT: I wrote this code to create columns of Z scores and then a total z score at the end. The result was that the columns showed up but were populated with NaN. Explain why this happened and how to correct it.

_______

Second attempt at this was done after changing the team_avgs dataframe to per set to match the tutorial instead of taking the mean of each category

```{r}

just_zscore <- vbzscore %>%
select(team, killszscore, aceszscore, digszscore, defkillzscore, errorzscore, hit_perc_zscore, service_error_zscore, TotalZscore)

view(just_zscore)

#big_ten = c('Wisconsin Badgers', 'Maryland Terrapins, Terps', 'Michigan Wolverines', 'Rutgers Scarlet Knights', 'Indiana Hoosiers', 'Iowa Hawkeyes', 'Penn St. Nittany Lions', 'Minnesota Golden Gophers', 'Illinois Fighting Illini', 'Ohio St. Buckeyes', 'Northwestern Wildcats', 'Michigan St. Spartans', 'Purdue Boilermakers')

just_zscore %>% 
filter(team == "Maryland Terrapins, Terps")
```

#Question: Why do teams show up two or three times in this dataset? Is it becasue the teams dataset shows every game for every team? Wouldn't the group_by(team) function solve that?

It just seems like my group_by isnt working. Running attempt 1 gives me a z-score column but attempt 2 of team_avgs does not. 

Analysis
- When I used attempt 2, NC State was high up in the total z-scores which tells me they are really good. They are 12-2 this season but in 2021, they finished 16-13. Not the bst record. IT suggested that maybe these metrics I chose aren't the strongest to measure a team by.

WIthout properly figuring out the issue with the group_by(team), a comprehensive analysis cannot be performed. I've used ChatGPT and Google to try and figuee out what the issue is as sghown in the copied prompts above. 

When I use attempw 2 on team_avgs, it geerates 340 rows which shows me that something at least worked in teh group_by. 

The code process seems to be correct for creating the z-scores and I understand how it should eb structured and what the steps are since I followed the tutorials. 
________________

z scores help you explain a team's excellence. 

_______________

Western Kentucky is a volleyball machine and its hitting percentage is well above average. The program has lost a combined 10 games over the last three seasons and made the second round of the NCAA in each of them.

It averages just over 13 kills per set and allows a touch over 11 and it has done so against some of the nation's top teams such as Kentucky and Notre Dame.

Many of its wins have come in straight sets showing that Western Kentucky has dominated its opponents. 



```{r}

rosters <- read_csv("https://raw.githubusercontent.com/Sports-Roster-Data/womens-volleyball/main/vb_rosters_2023_24.csv")

rosters_with_states <- rosters %>% 
  separate(Hometown, c("town", "state"), sep=",")
```



Dumbells and Lolly pop charts


```{r}
logs <- read_csv("https://dwillis.github.io/sports-data-files/ncaa_womens_volleyball_matchstats_2022.csv")

```

Looking at teams in the big ten and looking at their errors. This has the team, their errors and their opponents errors for last seaosn.
```{r}
big10 <- c("Nebraska Cornhuskers", "Iowa Hawkeyes", "Minnesota Golden Gophers", "Illinois Fighting Illini", "Northwestern Wildcats", "Wisconsin Badgers", "Indiana Hoosiers", "Purdue Boilermakers", "Ohio St. Buckeyes", "Michigan Wolverines", "Michigan St. Spartans", "Penn St. Nittany Lions", "Rutgers Scarlet Knights", "Maryland Terrapins, Terps")

errors <- logs |>
  filter(team %in% big10) |> 
  group_by(team) |> 
  summarise(
    total_errors = sum(errors), 
    opp_errors = sum(defensive_errors))
```


```{r}
ggplot() + 
  geom_dumbbell(
    data=errors, 
    aes(y=team, x=total_errors, xend=opp_errors),
    size = 1,
    colour = "grey",
    colour_x = "green",
    colour_xend = "red") +
  theme_minimal()
  
#xend is the end point of the of the dumbell
#issue is that we don't know which end is to which in the graph
#Ideally you want red on the right and green on the left
```

Making chart easier to see

```{r}
ggplot() + 
  geom_dumbbell(
    data=errors, 
    aes(y=team, x=total_errors, xend=opp_errors),
    size = 2,
    colour = "grey",
    colour_x = "green",
    colour_xend = "red") +
  theme_minimal()

```
Reordering to see the good stuff first.
Re ording to see the opponent errors first
This is the same as writing arrange(desc(opp_errors))
```{r}
ggplot() + 
  geom_dumbbell(
    data=errors, 
    aes(y=reorder(team, opp_errors), x=total_errors, xend=opp_errors),
    size = 2,
    colour = "grey",
    colour_x = "green",
    colour_xend = "red") +
  theme_minimal()


```

Lollipop Charts: They show magnatude and only show one dimention. 
 
```{r}
ggplot() + 
  geom_lollipop(
    data=errors, 
    aes(y=team, x=opp_errors), 
    horizontal = TRUE
    )

# you can do to line charts what you do with dumbell charts
```
```{r}
ggplot() + 
  geom_lollipop(
    data=errors, 
    aes(y=reorder(team, opp_errors), x=opp_errors), 
    horizontal = TRUE
    ) + theme_minimal() + 
  labs(title = "Maryland, Rutgers force more errors", y="team")

```

```{r}
md <- errors |> filter(team == "Maryland Terrapins, Terps")

ggplot() + 
  geom_lollipop(
    data=errors, 
    aes(y=reorder(team, opp_errors), x=opp_errors), 
    horizontal = TRUE
    ) + 
  geom_lollipop(
    data=md,
    aes(y=team, x=opp_errors),
    horizontal = TRUE,
    color = "red"
  ) + 
  theme_minimal() + 
  labs(title = "Maryland forced the most errors among Big Ten teams", y="team")

```



