```{r}
#install.packages("hoopR")
library(hoopR)
library(tidyverse)

options(scipen=999)
```
NYK team ID 1610612752
```{r}
lineups <-  nba_teamdashlineups(team_id = 1610612752)[[2]]
```

changing the col formats
```{r}

lineups <- lineups %>% 
  mutate(across(starts_with("GP"):ncol(lineups), as.numeric))

```

```{r}
lineups %>%
  filter(str_detect(GROUP_NAME, "Hartenstein")) %>% 
  filter(MIN != 0) %>% 
  select(GROUP_NAME, GP, MIN, PLUS_MINUS) %>% 
  arrange(desc(MIN))
  
```

Isaiah Hartenstein has been one of the iron men for the Knicks appearing in 82 games in 2022-23. When he's in the lineup the Knicks' highest plus/minus was a +19. Its lowest was a -12 but there has only been three games with that lineup as opposed to 10 with the highest plus/minus lineup.

Its a mix of bench players and starters there and so far they've played 84 minutes together which is nearly 50 minutes more than the second most-played lineup. My main point is that when Hartenstein is in the lineup, the Knicks. 



