```{r}

library(tidyverse)
library(janitor)
library(lubridate)
#install.packages("remotes")
#remotes::install_github("slu-openGIS/postmastr")
library(postmastr)
library(usdata)
library(rvest)
```

```{r}
# Loading data for the seaosns
logs24 <- read_csv("https://thescoop.org/sports-data-files/wbblogs24.csv")
logs18_24 <- read_csv("https://www.thescoop.org/sports-data-files/wbblogs1824.csv")

# removing the 2023-2024 from logs_18_24 to avoid duplicates when binding. Final product should have the first few games of 23-24 season.
logs18_24 <- logs18_24 %>% 
  filter(Season != "2023-2024")

logs <- bind_rows(logs18_24, logs24)

# Loading player data from 2022-2023
wbb23_24_rosters <- read_csv("data/wbb_rosters_2023_24.csv")
wbb22_23_rosters <- read_csv("data/wbb_rosters_2022_23.csv")
```

The countries that need to be cleaned are "The Netherlands and BOSNIA AND HERZEGOVINA"
NOTE: I replaced the 23_24 file with the one Derek updated: this is a link to the file. https://github.com/Sports-Roster-Data/womens-college-basketball/blob/main/wbb_rosters_2023_24.csv
```{r}
wbb22_23_rosters <- wbb22_23_rosters %>% 
  mutate(
    country_clean = case_when(
      country_clean == "THE NETHERLANDS" ~ "NETHERLANDS",
      country_clean == "BOSNIA AND HERZEGOVINA" ~ "BOSNIA & HERZEGOVINA",
      TRUE ~ as.character(country_clean)))
```

Creating power 5 list
```{r}
power_5 <- c("ACC", "Big 12", "Big Ten", "Pac-12", "SEC")
```

Filtering for non us players
```{r}
non_us_22_23 <- wbb22_23_rosters %>%
filter(country_clean != "USA", division == "I")
  
non_us_23_24 <-  wbb23_24_rosters %>% 
  filter(country_clean != "USA", division == "I")
```

What I want to do here is answer the obvious questions and generate dfs for each question showing the 22-23, the 23-24, diffs and percentage changes. 

createing teams datafraems
```{r}
#23-24
teams_with_int_players_23_24 <- non_us_23_24 %>% 
  group_by(ncaa_id, team, conference) %>% 
  summarize(int_players = n()) %>% 
  arrange(desc(int_players))

#22-23
teams_with_int_players_22_23 <- non_us_22_23 %>% 
  group_by(ncaa_id, team, conference) %>% 
  summarize(int_players = n()) %>% 
  arrange(desc(int_players))

# total players
total_player_23_24 <-  wbb23_24_rosters %>% 
  filter(division == "I") %>% 
  group_by(ncaa_id, team, conference) %>% 
  summarize(total_players = n()) %>% 
  arrange((team))

total_player_22_23 <- wbb22_23_rosters %>% 
  filter(division == "I") %>% 
  group_by(ncaa_id, team, conference) %>% 
  summarize(total_players = n()) %>% 
  arrange((team))

# this shows the missing team from 23-24 is Florida State
list <- anti_join (total_player_22_23, total_player_23_24, by= "team")
```

# need to figure out a new way to get a clean version of all players.

# I need to do the left joins in seperate years and then do the mutates for the NA values, then inner joins back together. 

```{r}
 
all_players23_24 <- left_join(total_player_23_24, teams_with_int_players_23_24, by = "ncaa_id")

all_players22_23 <- left_join(total_player_22_23, teams_with_int_players_22_23, by = "ncaa_id")

all_players <- inner_join(all_players23_24, all_players22_23, by = "ncaa_id")

```


```{r}
# cleaning teh all_players df
all_players <- all_players %>% 
  rename(team = team.x.x,
         conference = conference.x.x,
         total_23_24 = total_players.x,
         int_players_23_24 = int_players.x,
         total_22_23 = total_players.y,
         int_players_22_23 = int_players.y) %>% 
select(ncaa_id, team, conference, int_players_23_24, int_players_22_23, total_23_24, total_22_23, -team.y.x, -team.y.x, -team.x.y, -team.y.y)

# fixing NAs
all_players <- all_players %>%
  mutate(
    int_players_23_24 = case_when(
      is.na(int_players_23_24) ~ 0,
      TRUE ~ int_players_23_24),
    int_players_22_23 = case_when(
      is.na(int_players_22_23) ~ 0,
      TRUE ~ int_players_22_23))
```

adding in percentages
```{r}
# Percentage change doesn't show anything interesting here. Percentages does.

all_players <- all_players %>% 
  mutate(int_diff = int_players_23_24 - int_players_22_23,
         int_player_perc_23_24 = int_players_23_24/total_23_24*100,
         int_player_perc_22_23 = int_players_22_23/total_22_23*100)
```

player count categories
```{r}
all_players_categories <- all_players %>%
  mutate(
    category_22_23 = case_when(
        int_players_22_23 < 5 ~ "Less than 5",
        int_players_22_23 >= 5 & int_players_22_23 <= 10 ~ "Between 5 and 10 players",
        int_players_22_23 > 10 ~ "Greater than 10 players")) %>% 
  mutate(category_23_24 = case_when(
        int_players_23_24 < 5 ~ "Less than 5",
        int_players_23_24 >= 5 & int_players_23_24 <= 10 ~ "Between 5 and 10 players",
        int_players_23_24 > 10 ~ "Greater than 10 players"))
```

 Category counts
```{r}
categories_23_24 <- all_players_categories %>% 
  group_by(category_23_24 ) %>% 
  summarize(number_of_teams = n())

categories_22_23 <- all_players_categories %>% 
  group_by(category_22_23) %>% 
  summarize(number_of_teams = n())
```


------- Data Exploration Section -----------

Here I will answer the obvious outstanding questions and try to formulate a story narrative after it.

Which conference has most players
```{r}
# totals per conf
conf_counts23_24 <- wbb23_24_rosters %>% 
  filter(division == "I") %>% 
  group_by(conference) %>% 
  summarize(players_23_24 = n())
  
  conf_counts_22_23 <- wbb22_23_rosters %>% 
  filter(division == "I") %>% 
  group_by(conference) %>% 
  summarize(players_22_23 = n())
  
  total_conf <- inner_join(conf_counts23_24, conf_counts_22_23, by = "conference")
  
  #internationals
by_conf <- all_players %>% 
  group_by(conference) %>% 
  summarise(int_players_23_24 = sum(int_players_23_24),
            int_players_22_23 = sum(int_players_22_23))
#joining
all_conf <- inner_join(by_conf, total_conf, by="conference")

all_conf <- all_conf %>%  
  mutate(diff= int_players_23_24 - int_players_22_23,
         conf_perc_23_24 = int_players_23_24/sum(int_players_23_24)*100,
         conf_int_perc_22_23 = int_players_22_23/sum(int_players_22_23)*100)

# Filtering for the conference allows you to see a team's percentage of the conferences international players.
all_players %>%
  filter(conference == "AAC") %>% 
  summarise(total_int_players = sum(int_players_23_24)) %>% 
  mutate(team_perc_of_conf = total_int_players/sum(total_int_players)*100) %>% 
  arrange(desc(total_int_players))
```

which countries produce the most players
```{r}
#23-24
countries_23_24 <- non_us_23_24 %>% 
    group_by(country_clean) %>% 
  summarize(int_players = n()) %>% 
  arrange(desc(int_players))

#22-23
countries_22_23 <- non_us_22_23 %>% 
    group_by(country_clean) %>% 
  summarize(int_players = n()) %>% 
  arrange(desc(int_players))

# joining
total_country <- inner_join(countries_23_24, countries_22_23, by = "country_clean")

total_country <- total_country %>% 
  rename(int_players_23_24 = int_players.x,
         int_players_22_23 = int_players.y) %>% 
  mutate(diff = int_players_23_24 - int_players_22_23,
         perc_23_24 = int_players_23_24/sum(int_players_23_24)*100,
         perc_22_23 = int_players_22_23/sum(int_players_22_23)*100,
  perc_change = ((int_players_23_24 - int_players_22_23) / int_players_22_23) * 100)
```


--------- Second Presentation Progress Update Nov. 18, 2023 -------

The central question I want to answer is do teams with higher amounts of international players preform better than those with fewer international players. What I've done so far is create dfs to answer the obvious questions of "Which teams, conferences, countries, have the most international players in Division I basketball?" I also generated percentages of non-us players on each team and I want to make that a focal point of my analysis too. The teams with majority international players are ones I'd like to focus on more in-depth. These could be the teams with five or more international players, which, according to the code I wrote, is 52 teams. I chose to break it up by 15 and 10 because most teams have between 12 to 15 total players on their roster.

My next task will be to create comparisons to the previous year's data 2022-2023 to see if the number of international players grew or not? If so, by how much. Essentially, I'm running all the above codes to answer those same questions but with last season's data. I'll need to be intentional and careful in my joins so that I know what is what. Then, my task will be to tackle how to measure the performances of the teams. Another question would be do I also do a player analysis? Say I get player stats from the `wehoops` database. How would I mark certain players as internationals. I have not yet looked at this so I don't know if it gives player bio information along with stats.

The issues I've run into are about thinking beyond the obvious questions. I know the ways I can expand this, but figuring out which would best serve my analysis and get me closer to the central answer of my question is where one issue lies. This is something we could talk through to see if I'm missing any of the obvious questions I need to answer that could point me in the right direction of what type of expansion would best serve me.


_______ Progress after Second Update ---------

I updated the countries in 22-23 to match 23-24. There were only two that needed to be changed. This can be the section were I see if the teams with more international players do better than those who don't.

# I need the updated logs to include more games from this season

South Florida has the most international players in 23-24 so let's do a dive into them over the last season to now.

```{r}
# South Florida last season and this season
sf_22_24 <- logs %>% 
  filter(TeamFull == "South Florida Bulls", Season >= "2022-2023") 

```
z scores
```{r}
#averages
# filtering for 22-24 because that is all I have data on
team_avgs <- logs |>
  filter(Date >= "2022-10-01") %>% 
  group_by(Conference, Team) |> 
  summarise(
    FGAvg = mean(TeamFGPCT), 
    ReboundAvg = mean(TeamTotalRebounds), 
    OppFGAvg = mean(OpponentFGPCT),
    OffRebAvg = mean(OpponentTotalRebounds)
    ) %>% 
  rename(team = Team)

#Z scores
teamzscore <- team_avgs |> 
  mutate(
    FGzscore = as.numeric(scale(FGAvg, center = TRUE, scale = TRUE)),
    RebZscore = as.numeric(scale(ReboundAvg, center = TRUE, scale = TRUE)),
    OppZscore = as.numeric(scale(OppFGAvg, center = TRUE, scale = TRUE)) * -1,
    OppRebZScore = as.numeric(scale(OffRebAvg, center = TRUE, scale = TRUE)) * -1,
    TotalZscore = FGzscore + RebZscore + OppZscore + OppRebZScore
  )
```

joining with team ids
```{r}
#adding the number of international players

zscore_with_players <- left_join(teamzscore, all_players, by = "team") 

# the join shows 379 teams. but all players has 359 even though it should be 360.

#unique team ids
teams_id <- wbb23_24_rosters %>% 
  distinct(team, ncaa_id)

# generate unique teams from logs

unique_teams <- logs %>% 
  distinct(Team) %>% 
  rename(team = Team)

# join the names to teams_id
teams_and_ids <- left_join(unique_teams, teams_id, by ="team")

# populate the NA ncaa_ids using mutate

z_scores_and_ids <- left_join(zscore_with_players, teams_id, "ncaa_id" )
```


```{r}
# 63 teams that don't have ids
na_teams <- z_scores_and_ids %>% 
  filter(is.na(ncaa_id)) %>%
  select(team.x, ncaa_id)
  
# need to do a big case when to populate the missing ids.
  
teams_and_ids <- teams_and_ids %>%
  mutate(ncaa_id = case_when(
    team == "La Salle" ~ "340",
    team == "Loyola (IL)" ~ "371",
    team == "Saint Joseph's" ~ "606",
    team == "Jacksonville State" ~ "315",
    team == "UTSA" ~ "706",
    team == "Miami (FL)" ~ "415",
    team == "Albany (NY)" ~ "14",
    team == "Maryland-Baltimore County" ~ "391",
    team == "Massachusetts-Lowell" ~ "368",
    team == "NJIT" ~ "471",
    team == "TCU" ~ "698",
    team == "Connecticut" ~ "164",
    team == "St. John's (NY)" ~ "603",
    team == "Idaho State" ~ "294",
    team == "Presbyterian" ~ "1320",
    team == "South Carolina Upstate" ~ "10411",
    team == "Cal State Northridge" ~ "101",
    team == "North Carolina A&T" ~ "488",
    team == "UNC Wilmington" ~ "460",
    team == "Florida International" ~ "231",
    team == "Jacksonville State" ~ "315",
    team == "Sam Houston" ~ "624",
    team == "UTEP" ~ "704",
    team == "UTSA" ~ "706",
    team == "Western Kentucky" ~ "772",
    team == "Purdue Fort Wayne" ~ "308",
    team == "Wright State" ~ "810",
    team == "Youngstown State" ~ "817",
    team == "Bowling Green State" ~ "71",
    team == "Miami (OH)" ~ "414",
    team == "Morgan State" ~ "446",
    team == "South Carolina State" ~ "647",
    team == "Missouri State" ~ "669",
    team == "Southern Illinois" ~ "659",
    team == "Colorado State" ~ "156",
    team == "Nevada-Las Vegas" ~ "466",
    team == "Utah State" ~ "731",
    team == "Central Connecticut State" ~ "127",
    team == "Le Moyne" ~ "349",
    team == "Long Island University" ~ "361",
    team == "St. Francis (NY)" ~ "599",
    team == "Lindenwood" ~ "30136",
    team == "Little Rock" ~ "32",
    team == "SIU Edwardsville" ~ "660",
    team == "Southern Indiana" ~ "661",
    team == "Tennessee State" ~ "691",
    team == "Tennessee-Martin" ~ "695",
    team == "Southern California" ~ "657",
    team == "Boston University" ~ "68",
    team == "Loyola (MD)" ~ "369",
    team == "Louisiana State" ~ "365",
    team == "Jackson State" ~ "314",
    team == "Prairie View" ~ "553",
    team == "Texas A&M-Commerce" ~ "199",
    team == "Texas A&M-Corpus Christi" ~ "26172",
    team == "St. Thomas" ~ "620",
    team == "Arkansas State" ~ "30",
    team == "Louisiana-Monroe" ~ "498",
    team == "Southern Mississippi" ~ "664",
    team == "Tarleton State" ~ "1395",
    team == "Texas-Rio Grande Valley" ~ "536",
    team == "UT Arlington" ~ "702",
    team == "Pacific" ~ "2751",
  TRUE ~ as.character(ncaa_id)
  ))
```


```{r}
# Now, you can perform the inner join
#Remember what we talked about: populate the logs so that the ids are all in there, because everything proceeds from that.
# You’ll need to think about how to do that with two dataframes of IDs (the ones that match and the ones that you had to fix).



```
