```{r}

library(tidyverse)
library(janitor)
library(lubridate)
#install.packages("remotes")
#remotes::install_github("slu-openGIS/postmastr")
library(postmastr)
library(usdata)
library(rvest)
library(ggplot2)
library(ggalt)
#install.packages("ggtext")
library(ggtext)
library(htmltools)
```

```{r}
# Loading data for the seaosns
logs24 <- read_csv("https://thescoop.org/sports-data-files/wbblogs24.csv")
logs18_24 <- read_csv("https://www.thescoop.org/sports-data-files/wbblogs1824.csv")

# removing the 2023-2024 from logs_18_24 to avoid duplicates when binding. Final product should have the first few games of 23-24 season.
logs18_24 <- logs18_24 %>% 
  filter(Season != "2023-2024")

logs <- bind_rows(logs18_24, logs24)

# Loading player data from 2022-2023
wbb23_24_rosters <- read_csv("data/wbb_rosters_2023_24.csv")
wbb22_23_rosters <- read_csv("data/wbb_rosters_2022_23.csv")
```

The countries that need to be cleaned are "The Netherlands and BOSNIA AND HERZEGOVINA" NOTE: I replaced the 23_24 file with the one Derek updated: this is a link to the file. <https://github.com/Sports-Roster-Data/womens-college-basketball/blob/main/wbb_rosters_2023_24.csv>

```{r}
wbb22_23_rosters <- wbb22_23_rosters %>% 
  mutate(
    country_clean = case_when(
      country_clean == "THE NETHERLANDS" ~ "NETHERLANDS",
      country_clean == "BOSNIA AND HERZEGOVINA" ~ "BOSNIA & HERZEGOVINA",
      TRUE ~ as.character(country_clean)))
```

Creating power 5 list

```{r}
power_5 <- c("ACC", "Big 12", "Big Ten", "Pac-12", "SEC")
```

Filtering for non us players

```{r}
non_us_22_23 <- wbb22_23_rosters %>%
filter(country_clean != "USA", division == "I")
  
non_us_23_24 <-  wbb23_24_rosters %>% 
  filter(country_clean != "USA", division == "I")
```

What I want to do here is answer the obvious questions and generate dfs for each question showing the 22-23, the 23-24, diffs and percentage changes.

createing teams datafraems

```{r}
#23-24
teams_with_int_players_23_24 <- non_us_23_24 %>% 
  group_by(ncaa_id, team, conference) %>% 
  summarize(int_players = n()) %>% 
  arrange(desc(int_players))

#22-23
teams_with_int_players_22_23 <- non_us_22_23 %>% 
  group_by(ncaa_id, team, conference) %>% 
  summarize(int_players = n()) %>% 
  arrange(desc(int_players))

# total players
total_player_23_24 <-  wbb23_24_rosters %>% 
  filter(division == "I") %>% 
  group_by(ncaa_id, team, conference) %>% 
  summarize(total_players = n()) %>% 
  arrange((team))

total_player_22_23 <- wbb22_23_rosters %>% 
  filter(division == "I") %>% 
  group_by(ncaa_id, team, conference) %>% 
  summarize(total_players = n()) %>% 
  arrange((team))

# this shows the missing team from 23-24 is Florida State
list <- anti_join (total_player_22_23, total_player_23_24, by= "team")
```

# need to figure out a new way to get a clean version of all players.

# I need to do the left joins in seperate years and then do the mutates for the NA values, then inner joins back together.

```{r}
 
all_players23_24 <- left_join(total_player_23_24, teams_with_int_players_23_24, by = "ncaa_id")

all_players22_23 <- left_join(total_player_22_23, teams_with_int_players_22_23, by = "ncaa_id")

all_players <- inner_join(all_players23_24, all_players22_23, by = "ncaa_id")

```

```{r}
# cleaning teh all_players df
all_players <- all_players %>% 
  rename(team = team.x.x,
         conference = conference.x.x,
         total_23_24 = total_players.x,
         int_players_23_24 = int_players.x,
         total_22_23 = total_players.y,
         int_players_22_23 = int_players.y) %>% 
select(ncaa_id, team, conference, int_players_23_24, int_players_22_23, total_23_24, total_22_23, -team.y.x, -team.y.y, -team.y.y)

# fixing NAs
all_players <- all_players %>%
  mutate(
    int_players_23_24 = case_when(
      is.na(int_players_23_24) ~ 0,
      TRUE ~ int_players_23_24),
    int_players_22_23 = case_when(
      is.na(int_players_22_23) ~ 0,
      TRUE ~ int_players_22_23))
```

adding in percentages

```{r}
# Percentage change doesn't show anything interesting here. Percentages does.

all_players <- all_players %>% 
  mutate(int_diff = int_players_23_24 - int_players_22_23,
         int_player_perc_23_24 = int_players_23_24/total_23_24*100,
         int_player_perc_22_23 = int_players_22_23/total_22_23*100)
```

player count categories

```{r}
all_players_categories <- all_players %>%
  mutate(
    category_22_23 = case_when(
        int_players_22_23 < 5 ~ "Less than 5",
        int_players_22_23 >= 5 & int_players_22_23 <= 10 ~ "Between 5 and 10 players",
        int_players_22_23 > 10 ~ "Greater than 10 players")) %>% 
  mutate(category_23_24 = case_when(
        int_players_23_24 < 5 ~ "Less than 5",
        int_players_23_24 >= 5 & int_players_23_24 <= 10 ~ "Between 5 and 10 players",
        int_players_23_24 > 10 ~ "Greater than 10 players"))
```

Category counts

```{r}
categories_23_24 <- all_players_categories %>% 
  group_by(category_23_24 ) %>% 
  summarize(number_of_teams = n())

categories_22_23 <- all_players_categories %>% 
  group_by(category_22_23) %>% 
  summarize(number_of_teams = n())
```

------- Data Exploration Section -----------

Here I will answer the obvious outstanding questions and try to formulate a story narrative after it.

Which conference has most players

```{r}
# totals per conf
conf_counts23_24 <- non_us_23_24 %>% 
  filter(division == "I") %>% 
  group_by(conference) %>% 
  summarize(players_23_24 = n())
  
  conf_counts_22_23 <- non_us_22_23 %>% 
  filter(division == "I") %>% 
  group_by(conference) %>% 
  summarize(players_22_23 = n())
  
  total_conf <- inner_join(conf_counts23_24, conf_counts_22_23, by = "conference")
  
  #internationals
by_conf <- all_players %>% 
  group_by(conference) %>% 
  summarise(int_players_23_24 = sum(int_players_23_24),
            int_players_22_23 = sum(int_players_22_23))
#joining
all_conf <- inner_join(by_conf, total_conf, by="conference")

all_conf <- all_conf %>%  
  mutate(diff= int_players_23_24 - int_players_22_23,
         conf_perc_23_24 = int_players_23_24/sum(int_players_23_24)*100,
         conf_int_perc_22_23 = int_players_22_23/sum(int_players_22_23)*100)

# Filtering for the conference allows you to see a team's percentage of the conferences international players.
all_players %>%
  filter(conference == "AAC") %>% 
  summarise(total_int_players = sum(int_players_23_24)) %>% 
  mutate(team_perc_of_conf = total_int_players/sum(total_int_players)*100) %>% 
  arrange(desc(total_int_players))
```

which countries produce the most players

```{r}
#23-24
countries_23_24 <- non_us_23_24 %>% 
    group_by(country_clean) %>% 
  summarize(int_players = n()) %>% 
  arrange(desc(int_players))

#22-23
countries_22_23 <- non_us_22_23 %>% 
    group_by(country_clean) %>% 
  summarize(int_players = n()) %>% 
  arrange(desc(int_players))

# joining
total_country <- inner_join(countries_23_24, countries_22_23, by = "country_clean")

total_country <- total_country %>% 
  rename(int_players_23_24 = int_players.x,
         int_players_22_23 = int_players.y) %>% 
  mutate(diff = int_players_23_24 - int_players_22_23,
         perc_23_24 = int_players_23_24/sum(int_players_23_24)*100,
         perc_22_23 = int_players_22_23/sum(int_players_22_23)*100,
  perc_change = ((int_players_23_24 - int_players_22_23) / int_players_22_23) * 100)

# per team
non_us_22_23 %>% 
  filter(team == "South Florida") %>% 
  group_by(country_clean) %>% 
  summarize(int_players = n()) %>% 
  arrange(desc(int_players))


```

--------- Second Presentation Progress Update Nov. 18, 2023 -------

The central question I want to answer is do teams with higher amounts of international players preform better than those with fewer international players. What I've done so far is create dfs to answer the obvious questions of "Which teams, conferences, countries, have the most international players in Division I basketball?" I also generated percentages of non-us players on each team and I want to make that a focal point of my analysis too. The teams with majority international players are ones I'd like to focus on more in-depth. These could be the teams with five or more international players, which, according to the code I wrote, is 52 teams. I chose to break it up by 15 and 10 because most teams have between 12 to 15 total players on their roster.

My next task will be to create comparisons to the previous year's data 2022-2023 to see if the number of international players grew or not? If so, by how much. Essentially, I'm running all the above codes to answer those same questions but with last season's data. I'll need to be intentional and careful in my joins so that I know what is what. Then, my task will be to tackle how to measure the performances of the teams. Another question would be do I also do a player analysis? Say I get player stats from the `wehoops` database. How would I mark certain players as internationals. I have not yet looked at this so I don't know if it gives player bio information along with stats.

The issues I've run into are about thinking beyond the obvious questions. I know the ways I can expand this, but figuring out which would best serve my analysis and get me closer to the central answer of my question is where one issue lies. This is something we could talk through to see if I'm missing any of the obvious questions I need to answer that could point me in the right direction of what type of expansion would best serve me.

------ Progress after Second Update ---------

I updated the countries in 22-23 to match 23-24. There were only two that needed to be changed. This can be the section were I see if the teams with more international players do better than those who don't.

# I need the updated logs to include more games from this season

South Florida has the most international players in 23-24 so let's do a dive into them over the last season to now.

```{r}
# South Florida last season and this season
sf_22_24 <- logs %>% 
  filter(TeamFull == "South Florida Bulls", Season >= "2022-2023") 

```

z scores

```{r}
#averages
# filtering for 22-24 because that is all I have data on
team_avgs <- logs |>
  filter(Date >= "2022-10-01") %>% 
  group_by(Conference, Team) |> 
  summarise(
    FGAvg = mean(TeamFGPCT), 
    ReboundAvg = mean(TeamTotalRebounds), 
    OppFGAvg = mean(OpponentFGPCT),
    OffRebAvg = mean(OpponentTotalRebounds)
    ) %>% 
  rename(team = Team)

#Z scores
teamzscore <- team_avgs |> 
  mutate(
    FGzscore = as.numeric(scale(FGAvg, center = TRUE, scale = TRUE)),
    RebZscore = as.numeric(scale(ReboundAvg, center = TRUE, scale = TRUE)),
    OppZscore = as.numeric(scale(OppFGAvg, center = TRUE, scale = TRUE)) * -1,
    OppRebZScore = as.numeric(scale(OffRebAvg, center = TRUE, scale = TRUE)) * -1,
    TotalZscore = FGzscore + RebZscore + OppZscore + OppRebZScore
  )
```

joining with team ids

```{r}
#adding the number of international players

zscore_with_players <- left_join(teamzscore, all_players, by = "team") 

# the join shows 379 teams. but all players has 359 even though it should be 360.

#unique team ids
teams_id <- wbb23_24_rosters %>% 
  distinct(team, ncaa_id)

# generate unique teams from logs

unique_teams <- logs %>% 
  distinct(Team) %>% 
  rename(team = Team)

# join the names to teams_id
teams_and_ids <- inner_join(unique_teams, teams_id, by = "team")

teams_no_ids <- anti_join(unique_teams, teams_id, by = "team")


# populate the NA ncaa_ids using mutate

z_scores_and_ids <- left_join(zscore_with_players, teams_id, "ncaa_id" )
```

```{r}
# 63 teams that don't have ids
na_teams <- z_scores_and_ids %>% 
  filter(is.na(ncaa_id)) %>%
  select(team.x, ncaa_id)
  
# need to do a big case when to populate the missing ids.
  
teams_no_ids <- teams_no_ids %>%
  mutate(ncaa_id = case_when(
    team == "La Salle" ~ "340",
    team == "Loyola (IL)" ~ "371",
    team == "Saint Joseph's" ~ "606",
    team == "Jacksonville State" ~ "315",
    team == "UTSA" ~ "706",
    team == "Miami (FL)" ~ "415",
    team == "Albany (NY)" ~ "14",
    team == "Maryland-Baltimore County" ~ "391",
    team == "Massachusetts-Lowell" ~ "368",
    team == "NJIT" ~ "471",
    team == "TCU" ~ "698",
    team == "Connecticut" ~ "164",
    team == "St. John's (NY)" ~ "603",
    team == "Idaho State" ~ "294",
    team == "Presbyterian" ~ "1320",
    team == "South Carolina Upstate" ~ "10411",
    team == "Cal State Northridge" ~ "101",
    team == "North Carolina A&T" ~ "488",
    team == "UNC Wilmington" ~ "460",
    team == "Florida International" ~ "231",
    team == "Jacksonville State" ~ "315",
    team == "Sam Houston" ~ "624",
    team == "UTEP" ~ "704",
    team == "UTSA" ~ "706",
    team == "Western Kentucky" ~ "772",
    team == "Purdue Fort Wayne" ~ "308",
    team == "Wright State" ~ "810",
    team == "Youngstown State" ~ "817",
    team == "Bowling Green State" ~ "71",
    team == "Miami (OH)" ~ "414",
    team == "Morgan State" ~ "446",
    team == "South Carolina State" ~ "647",
    team == "Missouri State" ~ "669",
    team == "Southern Illinois" ~ "659",
    team == "Colorado State" ~ "156",
    team == "Nevada-Las Vegas" ~ "466",
    team == "Utah State" ~ "731",
    team == "Central Connecticut State" ~ "127",
    team == "Le Moyne" ~ "349",
    team == "Long Island University" ~ "361",
    team == "St. Francis (NY)" ~ "599",
    team == "Lindenwood" ~ "30136",
    team == "Little Rock" ~ "32",
    team == "SIU Edwardsville" ~ "660",
    team == "Southern Indiana" ~ "661",
    team == "Tennessee State" ~ "691",
    team == "Tennessee-Martin" ~ "695",
    team == "Southern California" ~ "657",
    team == "Boston University" ~ "68",
    team == "Loyola (MD)" ~ "369",
    team == "Louisiana State" ~ "365",
    team == "Jackson State" ~ "314",
    team == "Prairie View" ~ "553",
    team == "Texas A&M-Commerce" ~ "199",
    team == "Texas A&M-Corpus Christi" ~ "26172",
    team == "St. Thomas" ~ "620",
    team == "Arkansas State" ~ "30",
    team == "Louisiana-Monroe" ~ "498",
    team == "Southern Mississippi" ~ "664",
    team == "Tarleton State" ~ "1395",
    team == "Texas-Rio Grande Valley" ~ "536",
    team == "UT Arlington" ~ "702",
    team == "Pacific" ~ "2751"
  )) %>% 
  mutate(ncaa_id = as.numeric(ncaa_id))

ids <- z_scores_and_ids %>% 
  select(team.x,  ncaa_id) %>% 
  rename(Team = team.x)

# now all of the schools have ids in them. 
```

```{r}
# Now, you can perform the inner join

#Remember what we talked about: populate the logs so that the ids are all in there, because everything proceeds from that.
# You’ll need to think about how to do that with two dataframes of IDs (the ones that match and the ones that you had to fix).

#the ones that dont match are na_teams. the ones that match are teams_id

logs_with_ids <- inner_join(logs, teams_and_ids, join_by(Team == team))

logs_with_matches <- inner_join(logs, teams_no_ids, join_by(Team == team))

all_logs <- bind_rows(logs_with_matches, logs_with_ids)



```

now i can do the z score stuff on all_logs



```{r}
new_team_avgs <- all_logs |>
  filter(Date >= "2022-10-01") %>% 
  group_by(Team, ncaa_id) |> 
  summarise(
    FGAvg = mean(TeamFGPCT), 
    ReboundAvg = mean(TeamTotalRebounds), 
    OppFGAvg = mean(OpponentFGPCT),
    OffRebAvg = mean(OpponentTotalRebounds)
    ) %>% 
  rename(team = Team)

FGAvg_mean <- mean(new_team_avgs$FGAvg, na.rm = TRUE)
FGAvg_sd <- sd(new_team_avgs$FGAvg, na.rm = TRUE)
ReboundAvg_mean <- mean(new_team_avgs$ReboundAvg, na.rm = TRUE)
ReboundAvg_sd <- sd(new_team_avgs$ReboundAvg, na.rm = TRUE)
OppFGAvg_mean <- mean(new_team_avgs$OppFGAvg, na.rm = TRUE)
OppFGAvg_sd <- sd(new_team_avgs$OppFGAvg, na.rm = TRUE)
OffRebAvg_mean <- mean(new_team_avgs$OffRebAvg, na.rm = TRUE)
OffRebAvg_sd <- sd(new_team_avgs$OffRebAvg, na.rm = TRUE)

# Apply z-score calculation in mutate
new_teamzscore <- new_team_avgs |> 
  mutate(
FGzscore = (FGAvg - FGAvg_mean) / FGAvg_sd,
RebZscore = (ReboundAvg - ReboundAvg_mean) / ReboundAvg_sd,
OppZscore = ((OppFGAvg - OppFGAvg_mean) / OppFGAvg_sd) * -1,
OppRebZScore = ((OffRebAvg - OffRebAvg_mean) / OffRebAvg_sd) * -1,
TotalZscore = FGzscore + RebZscore + OppZscore + OppRebZScore)  

# When I run this, it produces NaN = not a number. 
```

joining

```{r}
players_zscores <- inner_join(new_teamzscore, all_players, by = "ncaa_id")
```

----------- Third update Dec. 3, 2023 ---------------

In this update, I was able to join my international player data to my game logs which was a struggle. I began calculating team averages and then z-scores so I can see if teams with a certain number of international players did better than ones with fewer international players.

However, when I ran the code to calculate z-scores, it produced NaN (Not a number) and after some research, I found online that this could be because there was a divide by 0 somewhere in the data. Finding that is what would be my next step and actually getting the z scores is what is essential to my overall question. Fixing the NaN is something we could work on figuring out how to do or maybe you could help guide me in a direction so I could figure it out myself. Of course, in a story, I would write around saying the phrase z-score obviously but I need to see how team's like South Florida which has 12 international players fares in a z-score analysis. Another question I had in this process is what US state gets the most international players? I haven't written the code yet for this question but can do it after this update because it would be good to know.

Visualizing the data would be done using stacked bar charts and maps. There could be two bars for each team. One with the number of international players and anther with their total players. Another choice would be a heat map of the countries represented.

--------- Progress after third update ---------------


```{r}
subset <- players_zscores %>% 
  select(team.x, ncaa_id, conference, TotalZscore, int_players_23_24, int_players_22_23) %>%
  arrange(desc(TotalZscore))
```


---- position analysis---


Which country produces the most at each position.
Create a df where the columns are the countries and various positions in years cols ex `pg_23_24` and `pg_22_23` and then `pg_diff` `sg_diff` maybe ditch records with fewer than 2 positions.

use pivot wider?
```{r}
#23-24
position_by_country_23_24 <- non_us_23_24 %>% 
  drop_na(position_clean) %>% 
  group_by(country_clean, position_clean) %>% 
  summarize(number_of_players = n()) %>% 
  #filter(number_of_players >= 3) %>% 
  pivot_wider(names_from = position_clean, values_from = number_of_players) %>%
  mutate_all(~replace(., is.na(.), 0))

#22-23
position_by_country_22_23 <- non_us_22_23 %>% 
  drop_na(position_clean) %>% 
  group_by(country_clean, position_clean) %>% 
  summarize(number_of_players = n()) %>% 
  #filter(number_of_players >= 3) %>% 
  pivot_wider(names_from = position_clean, values_from = number_of_players) %>%
  mutate_all(~replace(., is.na(.), 0))

# joining and differences
# .x = 23-24 .y=22-23
total_position_by_country <- inner_join(position_by_country_23_24, position_by_country_22_23, by = "country_clean")

total_position_by_country <- total_position_by_country %>% 
  rename(C_23_24 = CENTER.x,
         C_22_23 = CENTER.y,
         G_23_24 = GUARD.x,
         G_22_23 = GUARD.y,
         F_23_24 = FORWARD.x,
         F_22_23 = FORWARD.y,
         G_F_23_24 = `GUARD/FORWARD.x`,
         G_F_22_23 = `GUARD/FORWARD.y`,
         F_C_23_24 = `FORWARD/CENTER.x`,
         F_C_22_23 = `FORWARD/CENTER.y`,
         W_23_24 = WING.x,
         W_22_23 = WING.y)


#differences 
total_position_by_country <- total_position_by_country %>% 
mutate(F_diff = F_23_24 - F_22_23,
       C_diff = C_23_24 - C_22_23,
       G_diff = G_23_24 - G_22_23,
       G_F_diff = G_F_23_24 - G_F_22_23)
```


```{r}
# positions with NO COUNTRIES 

position_23_24 <- non_us_23_24 %>% 
  drop_na(position_clean) %>% 
  group_by(position_clean) %>% 
  summarize(number_of_players = n()) %>% 
  #filter(number_of_players >= 3) %>% 
  #pivot_wider(names_from = position_clean, values_from = number_of_players) %>%
  mutate_all(~replace(., is.na(.), 0))

#22-23
position_22_23 <- non_us_22_23 %>% 
  drop_na(position_clean) %>% 
  group_by(position_clean) %>% 
  summarize(number_of_players = n()) %>% 
  #filter(number_of_players >= 3) %>% 
  #pivot_wider(names_from = position_clean, values_from = number_of_players) %>%
  mutate_all(~replace(., is.na(.), 0))

# joining and differences
# .x = 23-24 .y=22-23
total_position <- inner_join(position_23_24, position_22_23, by = "position_clean")


total_position <- total_position %>% 
  rename(players_23_24 = number_of_players.x,
         players_22_23 = number_of_players.y) %>% 
  mutate(position_diff = players_23_24 - players_22_23)
```


----- A cleaner country analysis -------

Final product would be a df with the col names as `country_clean` and the rows as each `team` and `ncaa_id` and the number of players each `team` has from each `country_clean`

```{r}
#23-24
 country_counts_by_conf_23_24 <- non_us_23_24 %>% 
  group_by(country_clean, conference) %>% 
  summarize(int_players = n()) %>% 
  arrange(desc(int_players)) %>% 
  pivot_wider(names_from = conference, values_from = int_players) %>% 
  mutate_all(~replace(., is.na(.), 0))
# Spain is the only country to have at least one player in each conference

#22-23
 country_counts_by_conf_22_23 <- non_us_22_23 %>% 
  group_by(country_clean, conference) %>% 
  summarize(int_players = n()) %>% 
  arrange(desc(int_players)) %>% 
  pivot_wider(names_from = conference, values_from = int_players) %>% 
  mutate_all(~replace(., is.na(.), 0))

# Big Ten is the only conference to not have a player from Spain in 2022-23
 
```


----------- Final analysis summary -------------

It is clear some teams are making intentional decisions concerning their recruiting practices. South Florida has 80% of its roster this year as non-us players and 16 rosters in Division I women's college basketball have rosters that are 50% or greater international players. The change from this season to last was quite surprising across conferences, positions and countries. I took a deep look into all of those and answered the obvious. surface-level questions while takibng an in-depth look at other factors such as if teams with greater numbers of international players performed better than those who had fewer.

Starting with the team breakdown in the `all_players` dataframe. This shows the amount of international players last year and this year, the difference and the percentages of each team for last year and this year. What is most surprising to me is how many teams increased in the amount of international players. Whether it was an increase of three or five, it shows that the international game is growing.

When I was at Rider, I always knew many MAAC teams had international players and it has the fifth most this season with 49. Only up three from last year but that shows some degree of consistency. If I looked at more years in the repo, I could have identified a more concrete trend but looking at this season and last was the true scope of my analysis. The Big Ten was 20th on the list with 22 international players this season, down six from the prior year. I created the `all_conf` dataframe to look at this data by grouping_by conference and then counting the number of players. The percentages weren't as fruitful as they were in the `all_players` dataframe.

The next thing I wanted to look at was the countries and in that section, I found that Spain, Canada and Australia produce the most non-us players. I created a codeblock where you can enter a team name and see its beakdown of where its players come from. It would be interesting to see if I could create this into a quarto doc or something publishable where readers would be able to filter by team and see the number of international players. I'm unsure of how to do that at the current time.

```{r}
non_us_23_24 %>% 
  filter(team == "Maryland") %>% 
  group_by(country_clean) %>% 
  summarize(int_players = n()) %>% 
  arrange(desc(int_players))
```
I found it incredibly interesting that Australia actually produced fewer players from last year to this year. It decreased by 23 players whereas Spain 22% more players in 2023-24 than the prior year. That's equal to 26 new players.

```{r}

country_subset <- total_country %>%
 head(10)

# making the chart
ggplot() + 
  geom_dumbbell(
    data=country_subset, 
    aes(y=reorder(country_clean, int_players_23_24), x=int_players_22_23, xend=int_players_23_24),
    size = 2,
    color = "grey",
    colour_x = "red",
    colour_xend ="blue") + 
  theme_minimal()+
  expand_limits(x = 200)+
  labs(
    title = "International women's college basketball players year over year",
    x = "Number of players",
    y = "Country",
    caption = "Source: The Sports Roster Data Project at the University of Maryland | Chart by Dylan Manfre",
    subtitle = "Two of the top three countries with most international players increased since last year.")
  
```


The hardest part of the project was joining the teams and their international player counts with the z-score analysis. What I learned from this was that joinign by `ncaa_id` is better than joining by `team` because each id is unique and teams change names and/or close down. Joining taught me to be mindful of the order of my joins in that the dataframe with more rows needs to go in front than the one with fewer rows. the setup is (x,y, by="col_name"). Unfortunatly, the z-score analysis did not produce anything too interesting. It did not show that teams with more international players do better overall than the ones that have fewer. The top teams based on the categories I looked at were teams that were in the NCAA tournament. Of the top 20 teams, only South Florida had more than 10 players. Aside from that, nobody had more than 4.

This leads me into another thing I looked at. Categories. I thought that the number of teams in 23-24 who had 5 to 10 international players was higher than 2022-23 by 13.

Of the teams with international players, I broke it down by position to see what that comparison looks like. I found the position breakdown year over year to be quite interesting in terms of the types of players that are produced from various countries. After reworking the data and making sure the columns were correct, It was guards and forwards that were the most dominant positions among international players. Of the 65 countries listed in the `total_position_by_country` dataframe, 80 guards were from Spain this season. I found that particular dataframe harder to visualize which is why I went with the one below. 

```{r}
write_csv(total_position, "data/total_position.csv")

html_code <- '<iframe src="https://datawrapper.dwcdn.net/sN2ZW/2/" width="100%" height="489" scrolling="no" frameborder="0" allowtransparency="true"></iframe>'

HTML(html_code)
```




```{r}
Testing this to seee where an update appears. 
```




















