---
title: "Final Project: Do teams with internatinal players perform better than those who don't"
author: "Dylan Manfre"
date: '2023-11-07'
format: 
  html:
    code-fold: true
---

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
#install.packages("remotes")
#remotes::install_github("slu-openGIS/postmastr")
library(postmastr)
library(usdata)
library(rvest)
```

```{r}
# Loading data for the seaosns
logs18_24 <- read_csv("https://www.thescoop.org/sports-data-files/wbblogs1824.csv")

# Loading player data from 2022-2023
wbb23_24_rosters <- read_csv("data/wbb_rosters_2023_24.csv")

```

--------- Data cleaning ---------



-------- Exploratory data section ----------

What do I want to accomplish: I want to see what teams have the most international players and how those teams have performed over recent years to answer the core question: Do teams with international players that contribute do better than ones who do not. 

Steps I need to take:
- Filter for every non-usa player. 
- Do a group_by() and summarize() to see which team has the most international players in the 2022-23 season. 
- See if any of them have transfered prior to the 2022-23 season and see where.
- Do an analysis of the teams with non-usa players in logs18_24 generating their season stats.

What I still need to do:
- Need to get player stats so I can see player stats who have actually contributed to their teams. 
- How do I get the individual player stats for the players who are internationals?
- Do I just do a team analysis based on the teams that have internationals?


filter for non-us players and counting by team
```{r}
# filtering for non-us players
non_us_players <- wbb23_24_rosters %>% 
  filter(country_clean != "USA", division == "I")

# counting teams with international players
teams_with_int_players <- non_us_players %>% 
  group_by(team, conference) %>% 
  summarize(int_playrs = n()) %>% 
  arrange(desc(int_playrs))

# where do international players come from
## Spain, Austrailia and Canada are the top three which makes sense. Spanish national team is very good.
countries <- non_us_players %>% 
    group_by(country_clean) %>% 
  summarize(int_playrs = n()) %>% 
  arrange(desc(int_playrs))


```


Here I will create a bolian mid-major column to see how many teams here are mid-majors vs. p5 and then count them.
```{r}
power_5 <- c("ACC", "Big 12", "Big Ten", "Pac-12", "SEC")

# Assuming your data frame is named 'your_data' and the column is 'conference'

teams_with_int_players <- teams_with_int_players %>%
  mutate(mid_major_p5 = ifelse(conference %in% power_5, "power_5", "mid_major"))

# 316 more mid-majors than power_5 schools have international players. 
teams_with_int_players %>% 
  group_by(mid_major_p5) %>% 
  summarize(count = n())

```

I want to create a list of teams and their total players. Then I want to join that to `teams_with_international_players`. Then I want to create a percentage of international players on each team. I want to do this becasue in a story or analysis, percentages are easier to comprehend.

```{r}
# total_players df
total_players <-  wbb23_24_rosters %>% 
  group_by(team, conference) %>% 
  summarize(total_players = n()) %>% 
  arrange((team))

#joining

all_players <- inner_join(teams_with_int_players, total_players, by="team")

all_players <- all_players %>% select(-conference.y) %>% rename(conference = conference.x)

# perc_int

all_players <- all_players %>% 
  mutate(int_perc = (int_playrs/total_players)*100) %>% 
  arrange(desc(int_perc))

```










