---
title: "Final Project: Do teams with internatinal players perform better than those who don't"
author: "Dylan Manfre"
date: '2023-11-07'
format: 
  html:
    code-fold: true
---

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
#install.packages("remotes")
#remotes::install_github("slu-openGIS/postmastr")
library(postmastr)
library(usdata)
library(rvest)
```

```{r}
# Loading data for the seaosns
logs18_24 <- read_csv("https://www.thescoop.org/sports-data-files/wbblogs1824.csv")

# Loading player data from 2022-2023
wbb23_24_rosters <- read_csv("data/wbb_rosters_2023_24.csv")
wbb22_23_rosters <- read_csv("data/wbb_rosters_2022_23.csv")

```

The countries that need to be cleaned are "The Netherlands and BOSNIA AND HERZEGOVINA"

```{r}

wbb22_23_rosters <- wbb22_23_rosters %>% 
  mutate(
    country_clean = case_when(
      country_clean == "THE NETHERLANDS" ~ "NETHERLANDS",
      country_clean == "BOSNIA AND HERZEGOVINA" ~ "BOSNIA & HERZEGOVINA",
      TRUE ~ as.character(country_clean)))

```


NOTE: I replaced the 23_24 file with the one Derek updated: this is a link to the file. https://github.com/Sports-Roster-Data/womens-college-basketball/blob/main/wbb_rosters_2023_24.csv

-------- Exploratory data section for 2023-2024 ----------

What do I want to accomplish: I want to see what teams have the most international players and how those teams have performed over recent years to answer the core question: Do teams with international players that contribute do better than ones who do not. 

Steps I need to take:
- Filter for every non-usa player. 
- Do a group_by() and summarize() to see which team has the most international players in the 2022-23 season. 
- See if any of them have transfered prior to the 2022-23 season and see where.
- Do an analysis of the teams with non-usa players in logs18_24 generating their season stats.

What I still need to do:
- Need to get player stats so I can see player stats who have actually contributed to their teams. 
- How do I get the individual player stats for the players who are internationals?
- Do I just do a team analysis based on the teams that have internationals?

filter for non-us players and counting by team
```{r}
# filtering for non-us players
non_us_players <- wbb23_24_rosters %>% 
  filter(country_clean != "USA", division == "I")

# counting teams with international players
teams_with_int_players_23_24 <- non_us_players %>% 
  group_by(team, conference) %>% 
  summarize(int_players = n()) %>% 
  arrange(desc(int_players))

# where do international players come from
## Spain, Austraulia and Canada are the top three which makes sense. Spanish national team is very good.
countries_23_24 <- non_us_players %>% 
    group_by(country_clean) %>% 
  summarize(int_players = n()) %>% 
  arrange(desc(int_players))
```

Here I will create a bolian mid-major column to see how many teams here are mid-majors vs. p5 and then count them.
```{r}
power_5 <- c("ACC", "Big 12", "Big Ten", "Pac-12", "SEC")

teams_with_int_players_23_24 <- teams_with_int_players_23_24 %>%
  mutate(mid_major_p5 = ifelse(conference %in% power_5, "power_5", "mid_major"))

# 316 more mid-majors than power_5 schools have international players. 
teams_with_int_players_23_24 %>% 
  group_by(mid_major_p5) %>% 
  summarize(count = n())

```

# Possible question: What percentage of the conference total are on each team. I want to wrtie a sentence that hypothetically says: "Team X has 30% of the Big Ten's international players."

# From there, I can dive into how those players impact said teams.

```{r}
# Big Ten has 22 players
by_conf <- teams_with_int_players_23_24 %>% 
  group_by(conference) %>% 
  summarise(int_players = sum(int_players)) %>% 
  mutate(conf_int_perc = int_players/sum(int_players)*100) %>% 
  arrange(desc(int_players))

# South Florida has 40% of the AAC's international players and the most in D1 with 12.
# Filtering for the confernece allowes you to see a team's percentage of the conferences international players.
teams_with_int_players_23_24 %>%
  filter(conference == "AAC") %>% 
  summarise(total_int_players = sum(int_players)) %>% 
  mutate(team_perc_of_conf = total_int_players/sum(total_int_players)*100) %>% 
  arrange(desc(total_int_players))
```


STORY SENTENCE: `28% of Division I teams have one international player on its roster in 2023-24 season.`

```{r}
# The overwhelming majority have one player. This is obvious and not really newsworthy.
# How many teams have each number of players? Example sentence: 4 teams have 9 players.

number_of_teams_perc <- teams_with_int_players_23_24 %>% 
  group_by(int_players) %>% 
  summarise(num_teams = n()) %>%
  mutate(number_of_teams_perc = num_teams/sum(num_teams)*100) %>% 
  arrange(desc(num_teams))
```

STORY SENTENCE: `18.3% of teams have at least 5 international players on their roster this season.`
```{r}
at_least_5 <- number_of_teams_perc %>% 
  #filter(int_players >=5) %>% 
  summarise(at_least_5 = 52/sum(num_teams)*100)
```


I want to create a list of teams and their total players. Then I want to join that to `teams_with_international_players_23_24`. Then I want to create a percentage of international players on each team. I want to do this because in a story or analysis, percentages are easier to comprehend.

```{r}
# total_players df
total_players <-  wbb23_24_rosters %>% 
  group_by(team, conference) %>% 
  summarize(total_players = n()) %>% 
  arrange((team))

#joining

all_players <- inner_join(teams_with_int_players_23_24, total_players, by="team")

all_players <- all_players %>% select(-conference.y) %>% rename(conference = conference.x)

# perc_int

all_players <- all_players %>% 
  mutate(int_perc = (int_players/total_players)*100) %>% 
  arrange(desc(int_perc))


```

# What was the average amount of international players on a roster in 2022-23?
# I want to get an average amount of players, then filter out any teams that are below that average. The ones with more players are more likely the ones who will contribute more to their program.

Do something like this to form categories of players: "Less than five" "more than five" 

```{r}

all_players_categories <- all_players %>%
  mutate(
    int_players_categories = case_when(
        int_players < 5 ~ "Less than 5",
        int_players >= 5 & int_players <= 10 ~ "Between 5 and 10 players",
        int_players > 10 ~ "Greater than 10 players",
      )
  )
head(all_players_categories)

all_players_categories %>% 
  group_by(int_players_categories) %>% 
  summarize(number_of_teams = n())
```

Where do players on certain teams come from?
Change the name of the team in the filter to see any team you want.
```{r}
non_us_players %>% 
  filter(team == "South Florida") %>% 
  group_by(country_clean) %>%
  summarise(count = n()) %>% 
  arrange(desc(count))
```




----------- Making comparisons to 2022-23 season ------------
```{r}
# Non_us_22_23 
non_us_22_23 <- wbb22_23_rosters %>% 
  filter(country_clean != "USA", division == "I")

head(non_us_22_23)
```

```{r}
# total roster size 22-23
roster_size_22_23 <- wbb22_23_rosters %>% 
  group_by(team, conference) %>% 
  summarize(total_players = n()) %>% 
  arrange((team))

# non-us-players and then joining with roster size
int_players_22_23 <- non_us_22_23 %>% 
  group_by(team, conference) %>% 
  summarize(int_players = n()) %>% 
  arrange(desc(int_players))

# joining
all_players_22_23 <- inner_join(int_players_22_23, roster_size_22_23, by="team")
head(all_players_22_23)

all_players_22_23 <- all_players_22_23 %>% select(-conference.y) %>% rename(conference = conference.x)

# perc_int

all_players_22_23 <- all_players_22_23 %>% 
  mutate(int_perc = (int_players/total_players)*100) %>% 
  arrange(desc(int_players))

# categories 
team_categories_22_23 <- all_players_22_23 %>%
  mutate(
    int_players_categories = case_when(
        int_players < 5 ~ "Less than 5",
        int_players >= 5 & int_players <= 10 ~ "Between 5 and 10 players",
        int_players > 10 ~ "Greater than 10 players",))

# by_conf_22_23

by_conf_22_23 <- int_players_22_23 %>% 
  group_by(conference) %>% 
  summarise(int_players = sum(int_players)) %>% 
  mutate(conf_int_perc = int_players/sum(int_players)*100) %>% 
  arrange(desc(int_players))

```

```{r}
# 22-23 number of teams with int_players_perc
int_players_22_23 %>% 
  group_by(int_players) %>% 
  summarise(num_teams = n()) %>%
  mutate(number_of_teams_perc = num_teams/sum(num_teams)*100) %>% 
  arrange(desc(num_teams))
```

------ SECTION ON JOINS AND DIFFERENCES -------

My goal here will be to join certain dfs together. I will be joining the ones that are most important to my analysis. Then, I will rename columns to label them as a 2022-23 season or 23-24 season. Then I will calculate the difference or percent change if necessary.

If a team or conference, significantly increases the number of international players, that can be newsworthy and helpful in the analysis.

`These dfs will be key to my visualizations when I am at that stage of the project.`

analysis: The Big Ten, Big 12 and Sun Belt each lost six players while the ASUN gained 10 international players form 22-23 to the 23-24 season.
```{r}
# by conference
total_by_conf <- inner_join(by_conf, by_conf_22_23, by = "conference")
total_by_conf <- total_by_conf %>% 
  rename(int_players_23_24 = int_players.x,
          int_players_22_23 = int_players.y,
           perc_2023_24 = conf_int_perc.x,
          perc_2022_23 = conf_int_perc.y) %>%
  mutate(int_diff = int_players_23_24 - int_players_22_23)
```

```{r}
# by team category
categories_23_24 <- all_players_categories %>% 
  group_by(int_players_categories) %>% 
  summarize(number_of_teams = n())

categories_22_23 <- team_categories_22_23 %>% 
  group_by(int_players_categories) %>% 
  summarize(number_of_teams = n())
```

Analysis: There were 13 more teams in 2023-24 with 5-10 international players than in 22-23. 

```{r}
total_by_category <- inner_join(categories_23_24, categories_22_23, by = "int_players_categories")

total_by_category <- total_by_category %>%
  rename(teams_23_24 = number_of_teams.x,
         teams_22_23 = number_of_teams.y) %>% 
  mutate(diff = teams_23_24 - teams_22_23,
        grand_total = teams_23_24 + teams_22_23,
        perc_23_24 = teams_23_24/sum(teams_23_24)*100,
        perc_22_23 = teams_22_23/sum(teams_22_23)*100,
        perc_change = (perc_23_24-perc_22_23)/(perc_22_23)*100)

head(total_by_category)
```

Analysis: `Austraulia had the third most players in 23-24 but its down 23 players from a seaosn ago, which is 18%. Spain on the other hand, gained 24 new players.`

```{r}
# Countries and joining
countries_22_23 <- non_us_22_23 %>% 
    group_by(country_clean) %>% 
  summarize(int_players = n()) %>% 
  arrange(desc(int_players))

# joining
total_country <- inner_join(countries_23_24, countries_22_23, by = "country_clean")

total_country <- total_country %>% 
  rename(int_players_23_24 = int_players.x,
         int_players_22_23 = int_players.y) %>% 
  mutate(diff = int_players_23_24 - int_players_22_23, 
  perc_change = ((int_players_23_24 - int_players_22_23) / int_players_22_23) * 100)
```

```{r}
# Teams and joining

teams_total <- inner_join(teams_with_int_players_23_24, int_players_22_23, by = "team")

teams_total <- teams_total %>% 
  select(-conference.y) %>% 
  rename(int_players_23_24 = int_players.x,
         int_players_22_23 = int_players.y,
         conference = conference.x) %>% 
   mutate(diff = int_players_23_24 - int_players_22_23, 
  perc_change = ((int_players_23_24 - int_players_22_23) / int_players_22_23) * 100)
```


_______ Progress after Second Update ---------

I updated the countries in 22-23 to match 23-24. There were only two that needed to be changed. This can be the section were I see if the teams with more international players do better than those who don't.

# I need the updated logs to include more games from this season

South Florida has the most international players in 23-24 so let's do a dive into them over the last season to now.
```{r}

sf_22_24 <- logs18_24 %>% 
  filter(TeamFull == "South Florida Bulls", Season >= "2022-2023") 

```
z scores
```{r}
#averages
team_avgs <- logs18_24 |> 
  filter(Season >= "2022-2023") %>% 
  group_by(Conference, Team) |> 
  summarise(
    FGAvg = mean(TeamFGPCT), 
    ReboundAvg = mean(TeamTotalRebounds), 
    OppFGAvg = mean(OpponentFGPCT),
    OffRebAvg = mean(OpponentTotalRebounds)
    ) %>% 
  rename(team = Team)

#Z scores
teamzscore <- team_avgs |> 
  mutate(
    FGzscore = as.numeric(scale(FGAvg, center = TRUE, scale = TRUE)),
    RebZscore = as.numeric(scale(ReboundAvg, center = TRUE, scale = TRUE)),
    OppZscore = as.numeric(scale(OppFGAvg, center = TRUE, scale = TRUE)) * -1,
    OppRebZScore = as.numeric(scale(OffRebAvg, center = TRUE, scale = TRUE)) * -1,
    TotalZscore = FGzscore + RebZscore + OppZscore + OppRebZScore
  ) 

#adding the number of international players

zscore_with_players <- left_join(teamzscore, teams_total, by = "team") 

# I think this join is removing some teams? I only have 219 rows when team avg shows 377 teams.

```


```{r}


```

