---
name: Dylan Manfre
title: Does having a good assist to turnover ratio really lead to wins?

---

```{r}
#libraries
library(tidyverse)
library(lubridate)
library(janitor)
library(ggplot2)
#install.packages("ggrepel")
library(ggrepel)
library(rvest)
```

```{r}
# loading data from Her Hoops Stats library
game_boxes_with_teams <- read_csv("game_boxes_with_teams.csv")
View(game_boxes_with_teams)
```

Establishing lists
```{r}
 acc_teams <- c(
  "Boston College", "Clemson", "Duke", "Florida St.", "Georgia Tech", "Louisville", "Miami (FL)", "North Carolina", "NC State", "Pittsburgh", "Syracuse", "Virginia", "Virginia Tech", "Wake Forest")

  sec_teams = c(
    "Alabama", "Arkansas", "Auburn", "Florida", "Georgia", "Kentucky",
    "LSU", "Mississippi St.", "Missouri", "Ole Miss", "South Carolina", "Tennessee", "Texas A&M", "Vanderbilt")

 Big12 = c(
    "Baylor", "Iowa State", "Kansas", "Kansas St.", "Oklahoma",
    "Oklahoma St.", "TCU", "Texas", "Texas Tech", "West Virginia", "BYU", "Cincinnati", "UCF", "Houston", "Iowa St.")

   Pac12 = c(
    "Arizona", "Arizona St.", "California", "Colorado", "Oregon",
    "Oregon St.", "Stanford", "UCLA", "Southern California", "Utah", "Washington", "Washington St."
  )
   
big_ten <- c("Indiana", "Maryland", "Michigan", "Michigan St.", "Minnesota", "Nebraska", "Northwestern", "Ohio St.", "Penn St.", "Purdue", "Rutgers", "Illinois", "Iowa", "Wisconsin")

power_5 <- c("ACC", "Big 12", "Big Ten", "Pac-12", "SEC")
```

Data cleaning
```{r}
#initial cleaning
game_boxes_with_teams <- game_boxes_with_teams %>%
  rename(
    team_ator = ast_tov_ratio,
    opp_ator = opp_ast_tov_ratio,
    win_loss = win) %>%
  filter(!is.na(team_ator) & !(home_team_name == "Northwestern (IA)" | away_team_name == "Northwestern (IA)")) %>%
  mutate(
    winner = ifelse(opp_pts > pts, away_team_name, home_team_name),
    loser = ifelse(pts > opp_pts, away_team_name, home_team_name),
    winning_conference = ifelse(opp_pts > pts, away_team_conference, home_team_conference),
    losing_conference = ifelse(pts > opp_pts, away_team_conference, home_team_conference),
    winning_ator = ifelse(winner == home_team_name, team_ator, opp_ator),
    losing_ator = ifelse(loser == away_team_name, opp_ator, team_ator),
    difference = pts - opp_pts)

# making sure mutations worked
winner_loser <- game_boxes_with_teams %>%
  select(home_team_name, home_team_conference, pts, team_ator, away_team_name, away_team_conference, opp_pts, opp_ator, difference, winner, winning_conference, winning_ator, loser, losing_conference, losing_ator)
```
Narrative starts below.

Fans who listened to Maryland women's basketball head coach Brenda Frese's interviews last season heard the veteran coach say the words assist to turnover ratio. A basketball metric which divides assists by turnovers and is indicative of a team's ball security. 

Maryland advanced to the Elite Eight last year on a 1.4 assist to turnover ratio which was eighth best among power 5 teams and fourth in the Big Ten conference To get that number, I needed to make sure I understood what each team's assist to turnover ratio looked like. In a large dataset such as this one, trimming unnecessary columns made the work a lot easier.

```{r}
#picking columns I want
ator_only <- game_boxes_with_teams %>%
  select(gamedate_east, home_team_name, home_team_conference, pts, away_team_name, away_team_conference, opp_pts, winner, loser, winning_conference, losing_conference, ast, tov, opp_ast, opp_tov, team_ator, opp_ator, difference, winning_ator, losing_ator)

```

```{r}

# Get a unique list of team names
teams <- unique(ator_only$winner)
```

```{r}

# Aiden and I worked on this for loop together. He is a wizard and extreamly helpful with lists. We walked through things step by step and he was helpful with encouraging me to write the code and use different chat GPT prompts.

season_ator_list <- c()

for(team in teams){
  team_away <- ator_only %>% 
  filter(away_team_name == team)

team_away_ast <- sum(team_away$opp_ast)
team_away_turn <- sum(team_away$opp_tov)
  
team_home <- ator_only %>% 
  filter(home_team_name == team)

team_home_ast <- sum(team_home$ast)
team_home_turn <- sum(team_home$tov)

team_ator <-  (team_away_ast + team_home_ast)/(team_away_turn + team_home_turn)

season_ator_list <- c(season_ator_list, team_ator)}

```

```{r}

season_ator <- data.frame(team = teams, season_ator = season_ator_list)

```

Now I have something to work with. ator_only, gives me the team, opponent, game result and both teams' ators as well as other stats I believer are important to the analysis. But we need a little more before we dive deep. We need to find out a team's total wins, their ators and opponents.

```{r}
#Total_wins
total_wins <- game_boxes_with_teams %>%
  group_by(winner, winning_conference) %>%
  summarise(
    total_wins = n()) %>%
  arrange(desc(total_wins)) %>% 
  rename(team = winner)

# joining the season_ator to total_wins
wins_ator <- total_wins %>%
  inner_join(season_ator, by = "team")

```
Earlier in the season, a majority of teams had lower ators in November and December. It seems like teams did better in February and March, on average, which makes sense because they want to play better heading into the tournament make a good case for a high seed in the tournament. 

```{r}
time_element <- ator_only %>% 
  mutate(month = month(gamedate_east)) %>% 
  group_by(month) %>%
  summarise(avg_ator = mean(winning_ator),
            avg_opp_ator = mean(losing_ator)) %>% 
  arrange(desc(avg_ator))
```

We can make an early model to see if there is a possible relationship between a good ator and wins now that we have all the necessary components. A scatter plot is the best way to demonstrate this. We can contextualize this a bit more now that things are set up. Let's look at Maryland and see how its compared against the Big Ten and it's fellow power 5 schools. Knowing how Power 5 teams performed seems ideal for comparing Maryland because it's a majority of its competition.

```{r}
p5_teams_ator <- wins_ator %>% 
  filter(winning_conference %in% power_5)

b10 <- wins_ator %>% 
filter(team %in% big_ten)
```

The plot shows an upward slope which means there's at least some relationship between wins and a positive ator. So maybe Frese is on to something. 
```{r}
b10_only <- ggplot() +
  geom_point(data = b10, aes(x = total_wins, y = season_ator)) +
  geom_smooth(data=wins_ator, aes(x=total_wins, y=season_ator), method="lm") +
   geom_text_repel(
    data=b10, aes(x=total_wins, y=season_ator, label=team))+
  labs(title = "Big Ten Wins agaisnt Average Assist to Turnover Ratio",
       subtitle = "Maryland is on the right side of the assist to turnover ratio.", 
       caption = "Source: Her Hoops Stats | Graphic by Dylan Manfre",
       x = "Total Wins",
       y = "Team Assist to Turnover Ratio") 

```
Ask Iowa women's basketball head coach Lisa Bluder, and she'd agree with Brenda Frese. However, her Hawkeyes also have the highest opponent ator in the conference. But, the Hawkeyes also got 31 wins separating them from everyone else. 
```{r}
#only power 5 ators
p5_conf <- ator_only %>%
  filter(Conference %in% power_5) %>% 
  group_by(Conference) %>% 
  summarise(count = n(), 
            team_ator = mean(team_ator))

```
It's a close race but teams in the Big Ten had the highest ator last season. The Pac-12 had the second highest but not by much. There are many talented guards in the Big 12 and this is something I can look further into later on. I'm curious if there would be any centers who had a particularly high ator or non-guards.

Reading the model below, we can see the low R-squared and residual standard of error means there is a relationship between the two factors. Doing a linear model was the best way to explain this relationship becasue it gets to what Frese claims: that having a good assist to turnover ratio leads to wins. Looking at the scatter plot, we ccan tell that Maryland is doing pretty good for itself in terms of proving Frese's theory that having a good ator leads to wins.
```{r}

model1 <- lm(total_wins ~ season_ator, data = wins_ator)
summary(model1)
```
But lets spread it out over the course of the season. Filtering on a team's home and away name will show its full schedule. Nine of Maryland's highest team_ator's resulted in wins. The only loss was to Iowa in early March on the road. Even then, Maryland still had an ator of a 2 which is pretty respectable in Frese's eyes.

```{r}
umd <- ator_only %>% 
  filter(home_team_name == "Maryland" | away_team_name == "Maryland") 


umd <- umd %>% 
  mutate(umd_ator = ifelse(umd$winner == "Maryland", umd$winning_ator, umd$losing_ator)) %>% 
  select(gamedate_east, home_team_name, pts, away_team_name, opp_pts, team_ator, opp_ator, winner, winning_ator, losing_ator, umd_ator)


ggplot() + geom_line(data=umd, aes(x=gamedate_east, y=umd_ator))


```

In this scatter plot, I thought the SEC was interesting comparison, because like the Big Ten, its teams are spread around, but Tennessee stood out to me because of its placement. SEC teams seem to have a higher win count but lower ator.

```{r}
sec <- wins_ator %>% 
filter(team %in% sec_teams)

b10vSEC <- ggplot() +
  geom_point(data = b10, aes(x = total_wins, y = season_ator), color = "red") +
  geom_point(data = sec, aes(x = total_wins, y = season_ator), color = "blue") +
  geom_smooth(data=wins_ator, aes(x=total_wins, y=season_ator), method="lm") +
  labs(title = "Teams with strong guard depth have higher Assist to Turnover Ratios",
       subtitle = "Big Ten teams listed in red, SEC listed in blue.",
       x = "Total Wins",
       y = "Assist to Turnover Ratio in wins") +
  geom_text_repel(data = b10, aes(x = total_wins, y = season_ator, label = team))+
  geom_text_repel(data = sec, aes(x = total_wins, y = season_ator, label = team))

```

Let's look into Tennessee.
```{r}
tenn <- ator_only %>% 
  filter(home_team_name == "Tennessee" | away_team_name == "Tennessee") %>% 
   mutate(tenn_ator = ifelse(tenn$winner == "Tennessee", tenn$winning_ator, tenn$losing_ator))
```

Its best ator was a game against UCF when it had an ator of a 4.0 which is huge. They had 24 assists that game. UCF only had 11 assists and 15 turnovers. That particular game displayed what Tennessee is able to do on defense and how it can capitalize off opponents mistakes. 

Tennessee's highest allowed ator was a mid-November against Indiana who finished with a 2.0 and a 12-point win. The line graph shows a majority of Tennessee's games have ended with an above average ator which means it is playing efficient basketball. It ended the year with 25 wins and a Sweet 16 appearance. 

Looking at how it's spread out over time, the Vols stayed pretty consistent 1 for most of the season November was a 1.21 but it went 4-8 that month. December was it's best month averaging a 1.9 ator.

```{r}
ggplot() + geom_line(data=tenn, aes(x=gamedate_east, y=tenn_ator))+
            geom_hline(data=p5_conf, aes(yintercept=1.167473 ,color="red"))+
  labs(title="Tennessee is taking care of the ball",
       subtitle="Vols finished with an above average ator in a majority of its games.",
       caption="Source: Her Hoops Stats | Graph by Dylan Manfre",
       x = "Month",
       y = "Winning ator",
       color = "SEC Average ATOR in wins")
```
I noticed Tennessee won 12-of-13 games from Dec. 6 to Jan. 22. The codeblock below shows that Tennessee had a 1.7 ator in those 12. wins. Not quite a 2:1 but still pretty good all things considered because the avg_ator of all winners during that stretch of time was 1.07.
```{r}
#isolating the win streak
tenn_streaking <- ator_only %>% 
    filter((home_team_name == "Tennessee" | away_team_name == "Tennessee") &
         (gamedate_east >= "2022-12-06" & gamedate_east <= "2023-01-22")) %>% 
  select(gamedate_east, home_team_name, away_team_name, winner, winning_ator, losing_ator) %>% 
  arrange((gamedate_east))

# all teams ator during that stretch
ator_only %>% 
  filter(gamedate_east >= "2022-12-06" & gamedate_east <= "2023-01-22") %>%
  summarise(avg_ator = mean(winning_ator))

```
I scraped Tennessee's stats page. Table 22 gave me what I wanted with assists sand turnovers and ator so I kept that one. The table surprises me. Seven players have over a 1.0 which is respectable but two players have over a 2.0 which is ideal. I asked some coaches I know what they'd consider a good ator and they said a 2:1 is ideal. Those two players on Tennessee are Jasmine Powell and Edie Darby. I feel you have to eliminate Darby because she only played in 10 games whereas Powell played in 37. Darby also only had 2 assists and 1 turnover in those games.
```{r}
#Tenn player scrapping. Copied from the 772 book.

tenn_url <- "https://utsports.com/sports/womens-basketball/stats/2022-23"
 
# read in the html and extract all the tables
tenn_results <- tenn_url |>
  read_html() |>
  html_table()

tenn_results <- tenn_results[[22]]

tenn_playrs <- tenn_results %>% 
  clean_names()
```

The SEC, as we can see with the graph below, has the worst ator out of the power 5 teams which I figured was interesting because Tennessee seems to be so good in that statistic. It has a 0.4 lead over Kentucky, the next closest team. The conference as a whole, has the second most wins out of all the power 5 conferences too which also interested me. It shows an opposing point of view to Frese's point that maybe ator's don't always lead to wins, other components do too.

The Big 12 is to the far left with a respectable ator, but the fewest wins.

```{r}
p5_plot <- ggplot() + 
  geom_point(
    data=p5_conf, 
    aes(x=count, y=avg_winning_ator))+
  geom_text_repel(
    data=p5_conf, 
    aes(x=count, y=avg_winning_ator, label=winning_conference))+
  labs(title="Does a good ATOR really lead to more wins?",
       subtitle="The Big 12 shows the short answer is, it depends.",
       caption="Source: Her Hoop Stats | Graph by Dylan Manfre",
       x = "Wins",
       y = "Average Team ator")
```

There's a few characters at play here that both prove Frese's point and also dispel it. Clearly, having a good assist to turnover ratio does lead to wins in most cases. Every coach has their one aspect that they like to harp on. ator is hers and for Maryland, it's worked. The veteran head coach hopes it'll work again this year as the Terps hope to make it back to the Elite Eight with a fresh crop of players.

```{r}

 lsu <- ator_only %>% 
  filter(home_team_name == "lsu" | away_team_name == "lsu") %>% 
   mutate(lsu_ator = ifelse(winner == "lsu", winning_ator, losing_ator)) %>% 
  select(gamedate_east, winner, loser, winning_ator, losing_ator, lsu_ator)


```



